Weak interspecific interactions in a sagebrush steppe: evidence from observations, models, and experiments

### 2017-01-12 removals/manuscript/removal_ms.tex
More small edits

\title{Weak interspecific interactions in a sagebrush steppe: Evidence from observational data, population models, and a removal experiment}

@@

\begin{enumerate}
\item A previous study of four dominant species from a sagebrush steppe community used historical data sets and multispecies population models to demonstrate weak interspecific interactions, implying very little competitive release. 
\item We conducted a removal experiment to test the model's prediction of little competitive release. We established new quadrats in the same location where the historical data were collected. We assigned half the new quadrats to a  perennial grass removal treatment and half to a treatment in which we removed the dominant shrub, \textit{Artemisia tripartita}. We modeled survival, growth and recruitment as functions of local neighborhood species composition along with an indicator variable for removal treatment. If our ``baseline" model, which accounts for local plant-plant interactions, accurately predicts responses to removal, then the fitted removal indicator effect should be non-signficant.
\item  For survival and recruitment, the removal treatment effects were never significantly positive, indicating that competitive release was not underestimated by the interspecific effects included in the baseline model. For  \textit{Poa secunda} recruitment, the removal treatment was actually negative, indicating that our baseline model overestimated competitive release.
\item For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item For three of the four species, individual based models and integral projections models based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected greater competitive release than the baseline models. 
\item \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, overall the results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that interspecific resource competition is not the primary factor determining the dynamics of this community.

@@

Stable coexistence requires that intraspecific interactions must limit populations more than interspecific interactions \citep{chesson_mechanisms_2000}.  Whenever we study populations of species that appear to be coexisting stably, such as populations that have co-occurred in close proximity for extended periods, we should expect intraspecific limitations to be stronger than interspecific limitations. However, we still might expect interspecific interactions to be strong and negative. In fact, interspecific competition is arguably the primary obsession of plant community ecologists. We motivate research on plant species coexistence by noting that all plant species compete for water, light, and a few mineral resources (e.g. \citealt{silvertown_plant_2004}). Resource competition plays a central role in both Tilman's mechanistic resource competition models \citep{tilman_resource_1982} and Grime's Competitor-Stress tolerator-Ruderal strategies \citep{grime_plant_1979}. Neutral theory \citep{hubbell_unified_2001} goes even further, assuming that interspecific competition is exactly as strong as intraspecific competition. Recent work has emphasized the role of interspecific facilitation in some plant communities \citep{he_global_2013,brooker_facilitation_2008}, but it is unclear whether this research has shaken the default assumption that plant species compete intensely.

@@

The U.S. Sheep Experiment Station (USSES) is located 9.6 km north of Dubois, Idaho (44.2\degree N, 112.1\degree W), 1500 m above sea level. During the period of data collection (1926 – 1957), mean annual precipitation was 270 mm and mean temperatures ranged from -8\degree C (January) to 21\degree C (July). The vegetation is dominated by the shrub, \textit{Artemisia tripartita}, and the C3  perennial bunchgrasses \textit{Pseudoroegneria spicata}, \textit{Hesperostipa comata}, and \textit{Poa secunda}. These four species, the focus of our models, accounted for over 70\% of basal cover (grasses) and 60\% of canopy cover (shrubs and forbs).

@@

\subsection*{Removal experiment}
In spring 2011, in a large exclosure containing six of the historical permanent quadrats, we selected locations for an additional 16 quadrats for the removal treatments. We used the following criteria to select the quadrat locations. Each plot had to have 5\% canopy cover of \textit{Artemisia tripartita} and at least 3\% combined basal cover of the three focal perennial grasses. We avoided plots falling on hill slopes, areas with greater than 20\% bare rock, or with over 10\% cover of the woody shrubs \textit{Purshia tridentata} or \textit{Amelanchier utahensis}. We mapped the quadrats in June, 2011 and then implemented the removal treatments in September, 2011. We randomly assigned 8 of the sixteen new quadrats to the sagebrush removal treatment, and the remaining 8 to the grass removal treatment. We cut sagebrush stems at ground level and applied a 5\% solution of Roundup Pro (Glyphosate, Monsanto) to the cut stems. In the grass removal plots, we painted the same herbicide solution on all perennial grasses. We removed plants 50 cm beyond the quadrat boundaries to minimize edge effects. The sagebrush removal was virtually 100\% effective; grass removals required additional herbicide applications in April and May of each year, with fewer grasses remaining each year.

@@

Formally, we modeled the survival probability, $S$, of genet $i$ in species $j$, group $g$, and removal treatment $h$, from time $t$ to $t+1$  as
\begin{equation}
\mbox{logit}(S_{ijgh,t}) = \gamma_{j,t}^S + \varphi_{jg}^S+  \chi_{jh}^S  + \beta_{j,t}^S u_{ij,t} +  
\left \langle \boldsymbol{\omega}_{j,t}^S, \boldsymbol{W}_{ij,t} \right \rangle 
\label{eqn:survReg}
\end{equation}
where $\gamma$ is a time-dependent intercept, $\varphi$ is the coefficient for the effect of spatial location group, $\chi$ is the removal treatment effect, $\beta$ is the coefficient that represents the effect of log genet size, $u$, on survival. $\boldsymbol{\omega}$ is a vector of interaction coefficients which determine the impact of crowding, $\boldsymbol{W}$, by each species on the focal species. The vector $\boldsymbol{W}$ includes crowding from the four dominant species,  \textit{A. tripartita}, \textit{P. spicata}, \textit{H. comata}, and \textit{Poa secunda}, as well as two covariates representing total crowding from (1) all other perennial grasses and shrubs, which were mapped as polygons, and (2) all forb species, which were mapped as points. 
$\left \langle \boldsymbol{x, y} \right \rangle$ denotes the inner product of vectors $\boldsymbol{x}$ and $\boldsymbol{y}$, 
calculated as \texttt{sum(x*y)} in R.

@@

To simulate changes in cover in each quadrat from year $t$ to year $t+1$, we initialized the IBM with the actual configuration of the four focal species (genet sizes and locations) observed in year $t$. We then projected the model forward one year, applying the random year effects corresponding to year $t$ and the spatial random effects corresponding to the quadrat of interest. Because we were interested in comparing model predictions to observations, and were not interested in the effects of demographic stochasticity, we used a deterministic version of the model (e.g., recruitment is the $\lambda$ of (\ref{eqn:recrProcessModel}), rather than a random draw from a Poisson distribution with a mean of $\lambda$).
We first applied the recruitment regression to determine the expected number of new recruits at time $t+1$, which we then multiplied by the mean size of recruits observed for the focal species to get expected cover of new recruits. We then used the survival regression to predict the survival probability of each genet, and the growth regression to predict the change in size of each genet. Expected cover at $t+1$ is expected area of new recruits, plus the sum of the survival probabilities multiplied by the expected sizes.

@@

Our vital rate models take into account the effects of local species composition and should therefore be able to predict the effect of removals on survival, growth and recruitment without any additional parameters. If the removal treatment effects that we added to these models are statistically significant, it would indicate an effect of removals that goes above and beyond the response predicted by a model based on data from the control plots. For the survival models, the 95\% confidence intervals on the removal effects overlapped zero for all four species (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRsurvival} to \ref{PSSPsurvival}), meaning that the survival probabilities of plants in removal plots can be predicted based on data from control plots. However, for the growth models, the 95\% confidence intervals on the removal effects were positive for \textit{Poa secunda}  and \textit{P. spicata}, meaning that individuals of these species grew more rapidly in removal plots than controls even after accounting for neighborhood interactions (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRgrowth} to \ref{PSSPgrowth}). In contrast, \textit{Poa secunda} had lower recruitment in removal than controls plots (upper limit of 95\% CI $<0$; Fig. \ref{fig:VitalRateTest}; Table \ref{table:recruitment}).

@@

The positive removal effects on \textit{Poa secunda}  and \textit{P. spicata} growth are particularly interesting, because they suggest that models based on observational data may underestimate competitive release. We conducted additional analyses to further explore removal effects on the growth of these two species. First, we classified all individuals of these species as located either inside or outside the limits of \textit{A. tripartita} canopies present in 2011 (before removal). If removal is releasing grasses from \textit{A. tripartita} competition, we would expect individuals located under the shrub canopies to experience a greater effect of removal. However, the location of individuals with respect to \textit{A. tripartita} canopies had no effect on growth of either species (Tables \ref{table:POSEgrowth-inARTR} and \ref{table:PSSPgrowth-inARTR}), meaning that individuals located outside \textit{A. tripartita} canopies responded to removals as much as individuals underneath (removed) canopies.

@@

Third, responses to removals might be poorly predicted by models based on observational data if the removals create local crowding outside the range of historical variation or if the removals break covariances in crowding among species. However, the distribution of crowding observed within control plots easily spans the crowding values observed in removal plots (Fig. \ref{fig:Wscatters}), and we failed to find evidence for strong interspecific covariances in crowding (\ref{suppMethods}).

Finally, we evaluated pre-existing differences in community composition in removal and control plots. In selecting removal plots, our criteria required candidate plots to include a certain amount of our target removal species. Because we adopted the historical plots as our controls, the same criteria were not applied (a control plot might not contain any \textit{A. tripartita}). To quantify resulting differences in pre-treatment (2011) composition, we compared neighborhood crowding at the individual plant level (the $\boldsymbol{W}$ of the growth regression) across treatments. \textit{A. tripartita} crowding and crowding by all subdominant shrub and grass species appeared greater than in the controls, whereas crowding by \textit{H. comata} was lower in removal plots than controls, though it was absent from most \textit{P. spicata} neighborhood in both treatments (Fig. \ref{fig:W-by-treatment}). These pre-existing differences in composition raise the possibility that plots selected for the removal treatments might differ from the control plots in some unobserved covariate that could influence plant performance.

@@

For three of the four species, we find little or no evidence that models based on observational data underestimate the intensity of interspecific competition. Removal treatment effects were never significant for \textit{A. tripartita} or \textit{H. comata}; our baseline model, which accounts for local neighborhood composition, adequately explained vital rate responses to  removal.  \textit{Poa secunda} showed mixed responses to \textit{Artemisia} removal: removal effects were not significant at the survival stage, were positive at the growth stage, and negative at the recruitment stage (where the estimated effect of \textit{Artemisia} on \textit{Poa} is positive). At the population level, models with and without removal effects projected similar patterns of short (Fig. \ref{fig:IBM1step}) and long-term (Fig. \ref{fig:IPMresults}) dynamics for \textit{H. comata} and \textit{A. tripartita}. For \textit{Poa secunda}, the model with removal effects predicted higher population growth over the short but not the long-term, perhaps because the short-term projections mostly reflect the positive effect of \textit{Artemisia} removal on growth, while the long-term equilibrium are also sensitive to the negative effects of removal on recruitment. Although the baseline model did not always successfully predict removal effects on vital rates, it does appear to successfully capture the population-level consequences of removals on these three species and did not severely underestimate the magnitude of competitive release.

@@

The small competitive release of many species in our system implies that each of the four dominant species we studied are limited by quite different factors. This low niche overlap could reflect strong resource partitioning (Tilman 1982), perhaps combined with variation in the times and locations at which species are taking up those resources. Species-specific responses to spatial and temporal variation in non-resource environmental variables can also weaken interspecific interactions \citep{chesson_mechanisms_2000}, though in a previous study we failed to find evidence of a temporal storage effect in this community \citep{adler_weak_2009}. Interactions with above- or belowground enemies can regulate populations and, given enough species-specificity in the effects of these enemies, decouple interspecific dynamics \citep{hersh_evaluating_2011,janzen_herbivores_1970,connell_role_1971}.  More likely, plants may alter the environment, and even resource availability, in ways that facilitate their neighbors. For example, shade cast by \textit{Artemisia} canopies may reduce evapotranspiration (e.g. \citealt{Barbier2008}). \textit{Artemisia} removal might offest this facilitative effect by reducing resource uptake, resulting in little net change in water balance following removal.  Unfortunately, we did not monitor soil moisture at the fine spatial and temporal resolution necessary to test this hypothesis.

Evidence for weak interspecific interactions and limited competitive release carries interesting implications. If most interspecific interactions are weak, then single-species models may be adequate for forecasting population responses to environmental change \citep{adler_forecasting_2012,kleinhesselink_indirect_2015}. Similarly, differences among species in relative abundance should reflect variation in the strength of intraspecific density dependence more than variation in the strength of interspecific interactions \citep{yenni_strong_2012}. In such a system, effective management and conservation requires understanding the sources and species-specificity of intraspecific density dependence. In fact, an experiment to test our model's skill at predicting release from intraspecific competition could be more important than the test of interspecific competitive release described here.

@@

\newpage
\bibliographystyle{Ecology}
\bibliography{IPMrefs}

>>>>

\title{Weak interspecific interactions in a sagebrush steppe: evidence from observations, models, and experiments}

@@

\begin{enumerate}
\item A previous study of four dominant species from a sagebrush steppe community used historical data sets and multispecies population models to demonstrate weak interspecific interactions. The models predict little competitive release following species removals.
\item We conducted a removal experiment to test this prediction. We established new quadrats in the same location where the historical data were collected. We assigned half the new quadrats to a  perennial grass removal treatment and half to a treatment in which we removed the dominant shrub, \textit{Artemisia tripartita}. We modeled survival, growth and recruitment as functions of local neighborhood species composition along with an indicator variable for removal treatment. If our ``baseline" model, which accounts for local plant-plant interactions, accurately predicts responses to removal, then the fitted removal indicator effect should be non-signficant.
\item  For survival and recruitment, the removal treatment effects were never significantly positive, indicating that competitive release was not underestimated by the interspecific effects included in the baseline model. For  \textit{Poa secunda} recruitment, the removal treatment was actually negative, indicating that our baseline model overestimated competitive release.
\item For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item For three species, individual based models and integral projections models based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected greater competitive release than the baseline models. 
\item \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, the experimental results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that interspecific resource competition is not the primary factor determining the dynamics of this community.

@@

Stable coexistence requires that intraspecific interactions must limit populations more than interspecific interactions \citep{chesson_mechanisms_2000}.  Whenever we study populations of species that appear to be coexisting stably, such as populations that have co-occurred in close proximity for extended periods, we should expect intraspecific limitations to be stronger than interspecific limitations. However, we still might expect interspecific interactions to be strong and negative. In fact, plant community ecologists often motivate research on coexistence by noting that all plant species compete for water, light, and a few mineral resources (e.g. \citealt{silvertown_plant_2004}). Resource competition plays a central role in both Tilman's mechanistic resource competition models \citep{tilman_resource_1982} and Grime's Competitor-Stress tolerator-Ruderal strategies \citep{grime_plant_1979}. Neutral theory \citep{hubbell_unified_2001} goes even further, assuming that interspecific competition is exactly as strong as intraspecific competition. A growing list of studies has emphasized the role of interspecific facilitation in some plant communities \citep{he_global_2013,brooker_facilitation_2008}, but it is unclear whether this research has overturned the default assumption that plants compete intensely with other species.

@@

The U.S. Sheep Experiment Station (USSES) is located 9.6 km north of Dubois, Idaho (44.2\degree N, 112.1\degree W), 1500 m above sea level. During the period of data collection (1926 \textendash 1957), mean annual precipitation was 270 mm and mean temperatures ranged from -8\degree C (January) to 21\degree C (July). The vegetation is dominated by the shrub, \textit{Artemisia tripartita}, and the C3  perennial bunchgrasses \textit{Pseudoroegneria spicata}, \textit{Hesperostipa comata}, and \textit{Poa secunda}. These four species, the focus of our models, accounted for over 70\% of basal cover (grasses) and 60\% of canopy cover (shrubs and forbs).

@@

\subsection*{Removal experiment}
In spring 2011, in a large exclosure containing six of the historical permanent quadrats, we established an additional 16 quadrats for the removal treatments. New quadrats were selected to be similar to historical plots, but also needed to have enough of the focal species present so that removal experiments could be conducted. We evaluated potential locations for the new quadrats near the original quadrats and then used only those locations with at least 5\% canopy cover of \textit{Artemisia tripartita} and at least 3\% combined basal cover of the three focal perennial grasses. We rejected locations falling on hill slopes, areas with greater than 20\% bare rock, or with over 10\% cover of the woody shrubs \textit{Purshia tridentata} or \textit{Amelanchier utahensis}. We mapped the new quadrats in June, 2011 and then implemented the removal treatments in September, 2011. We randomly assigned 8 of the sixteen new quadrats to the sagebrush removal treatment, and the remaining 8 to the grass removal treatment. We cut sagebrush stems at ground level and applied a 5\% solution of Roundup Pro (Glyphosate, Monsanto) to the cut stems. In the grass removal plots, we painted the same herbicide solution on all perennial grasses. We removed plants 50 cm beyond the quadrat boundaries to minimize edge effects. The sagebrush removal was virtually 100\% effective; grass removals required additional herbicide applications in April and May of each year, with fewer grasses remaining each year.

@@

Formally, we modeled the survival probability, $S$, of genet $i$ in species $j$, group $g$, and removal treatment $h$, from time $t$ to $t+1$  as
\begin{equation}
\mbox{logit}(S_{ijgh,t+1}) = \gamma_{j,t}^S + \varphi_{jg}^S+  \chi_{jh}^S  + \beta_{j,t}^S u_{ij,t} +  
\left \langle \boldsymbol{\omega}_{j,t}^S, \boldsymbol{W}_{ij,t} \right \rangle 
\label{eqn:survReg}
\end{equation}
where $\gamma$ is a time-dependent intercept, $\varphi$ is the coefficient for the effect of spatial location group, $\chi$ is the removal treatment effect, $\beta$ is the coefficient that represents the effect of log genet size, $u$, on survival. $\boldsymbol{\omega}$ is a vector of interaction coefficients which determine the impact of crowding, $\boldsymbol{W}$, by each species on the focal species. The vector $\boldsymbol{W}$ includes crowding from the four dominant species,  \textit{A. tripartita}, \textit{P. spicata}, \textit{H. comata}, and \textit{Poa secunda}, as well as two covariates representing total crowding from (1) all other perennial grasses and shrubs, which were mapped as polygons, and (2) all forb species, which were mapped as points. 
$\left \langle \boldsymbol{x, y} \right \rangle$ denotes the inner product of vectors $\boldsymbol{x}$ and $\boldsymbol{y}$, 
calculated as \texttt{sum(x*y)} in R.

@@

To simulate changes in cover in each quadrat from year $t$ to year $t+1$, we initialized the IBM with the actual configuration of the four focal species (genet sizes and locations) observed in year $t$. We then projected the model forward one year, applying the random year effects corresponding to year $t$ and the spatial random effects corresponding to the quadrat of interest. Because we were interested in comparing model predictions to observations, and were not interested in the effects of demographic stochasticity, we used a deterministic version of the model (e.g., recruitment is the $\lambda$ of (\ref{eqn:recrProcessModel}), rather than a random draw from a Poisson distribution with a mean of $\lambda$).
We first applied the recruitment regression to determine the expected number of new recruits at time $t+1$, which we then multiplied by the mean size of recruits observed for the focal species to get expected cover of new recruits. We then used the survival regression to predict the survival probability of each genet, and the growth regression to predict the change in size of each genet. Expected cover at $t+1$ is expected area of new recruits, plus the sum of the survival probabilities multiplied by the expected sizes. Because we do not carry these simulations forward beyond one time step, we do not need to assign locations to the new recruits.

@@

Our vital rate models take into account the effects of local species composition and should therefore be able to explain the effect of removals on survival, growth and recruitment without any additional parameters. If the removal treatment effects that we added to these models are statistically significant, it would indicate an effect of removals that goes above and beyond the response predicted by a model based on data from the control plots. For the survival models, the 95\% confidence intervals on the removal effects overlapped zero for all four species (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRsurvival} to \ref{PSSPsurvival}), meaning that the survival probabilities of plants in removal plots can be predicted based on data from control plots. However, for the growth models, the 95\% confidence intervals on the removal effects were positive for \textit{Poa secunda}  and \textit{P. spicata}, meaning that individuals of these species grew more rapidly in removal plots than controls even after accounting for neighborhood interactions (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRgrowth} to \ref{PSSPgrowth}). In contrast, \textit{Poa secunda} had lower recruitment in removal than controls plots (upper limit of 95\% CI $<0$; Fig. \ref{fig:VitalRateTest}; Table \ref{table:recruitment}). For the other species, we could not detect significant removal treatment effects on recruitment.

@@

The positive removal effects on \textit{Poa secunda}  and \textit{P. spicata} growth suggest that models based on observational data may underestimate competitive release. We conducted additional analyses to further explore removal effects on the growth of these two species. First, we classified all individuals of these species as located either inside or outside the limits of \textit{A. tripartita} canopies present in 2011 (before removal). If removal is releasing grasses from \textit{A. tripartita} competition, we would expect individuals located under the shrub canopies to experience a greater effect of removal. However, the location of individuals with respect to \textit{A. tripartita} canopies had no effect on growth of either species (Tables \ref{table:POSEgrowth-inARTR} and \ref{table:PSSPgrowth-inARTR}), meaning that individuals located outside \textit{A. tripartita} canopies responded to removals as much as individuals underneath (removed) canopies.

@@

Third, responses to removals might be poorly predicted by models based on observational data if the removals reduce plant densities (crowding) below the range of historical variation, or if the removals break covariances in plant densities among species. However, the distribution of crowding observed within control plots easily spans the crowding values observed in removal plots (Fig. \ref{fig:Wscatters}), and we failed to find evidence for strong interspecific covariances in crowding (\ref{suppMethods}).

Finally, we evaluated pre-existing differences in community composition in removal and control plots. In selecting removal plots, our criteria required candidate plots to include a certain amount of our target removal species. Because we adopted the historical plots as our controls, the same criteria were not applied (a control plot might not contain any \textit{A. tripartita}). To quantify resulting differences in pre-treatment (2011) composition, we compared neighborhood crowding at the individual plant level (the $\boldsymbol{W}$ of the growth regression) across treatments. \textit{A. tripartita} crowding and crowding by all subdominant shrub and grass species appeared greater than in the controls, whereas crowding by \textit{H. comata} was lower in removal plots than controls, though it was absent from most \textit{P. spicata} neighborhoods in both treatments (Fig. \ref{fig:W-by-treatment}). These pre-existing differences in composition raise the possibility that plots selected for the removal treatments might differ from the control plots in some unobserved covariate that could influence plant performance.

@@

For three of the four species, we find little or no evidence that models based on observational data underestimate the intensity of interspecific competition. After accounting for local neighborhood crowding, the removal treatment effects were never significant for \textit{A. tripartita} or \textit{H. comata}; our baseline model, which accounts for local neighborhood composition, adequately explained vital rate responses to  removal.  \textit{Poa secunda} showed mixed responses to \textit{Artemisia} removal: removal effects were not significant at the survival stage, were positive at the growth stage, and negative at the recruitment stage (where the estimated effect of \textit{Artemisia} on \textit{Poa} is positive). At the population level, models with and without removal effects projected similar patterns of short (Fig. \ref{fig:IBM1step}) and long-term (Fig. \ref{fig:IPMresults}) dynamics for \textit{H. comata} and \textit{A. tripartita}. For \textit{Poa secunda}, the model with removal effects predicted higher population growth over the short but not the long-term, perhaps because the short-term projections mostly reflect the positive effect of \textit{Artemisia} removal on growth, while the long-term equilibrium are also sensitive to the negative effects of removal on recruitment. Although the baseline model did not always successfully predict removal effects on vital rates, it does appear to successfully capture the population-level consequences of removals on these three species and did not severely underestimate the magnitude of competitive release.

@@

The small competitive release of many species in our system implies that each of the four dominant species we studied are limited by quite different factors. This low niche overlap could reflect strong resource partitioning (Tilman 1982), perhaps combined with variation in the times and locations at which species are taking up those resources. Species-specific responses to spatial and temporal variation in non-resource environmental variables can also weaken interspecific interactions \citep{chesson_mechanisms_2000}, though in a previous study we failed to find evidence of a temporal storage effect in this community \citep{adler_weak_2009}. Interactions with above- or belowground enemies can regulate populations and, given enough species-specificity in the effects of these enemies, decouple interspecific dynamics \citep{hersh_evaluating_2011,janzen_herbivores_1970,connell_role_1971}.  More likely, plants may alter the environment, and even resource availability, in ways that facilitate their neighbors. For example, shade cast by \textit{Artemisia} canopies may reduce evapotranspiration (e.g. \citealt{Barbier2008}). \textit{Artemisia} removal might offset this facilitative effect by reducing resource uptake, resulting in little net change in water balance following removal.  Unfortunately, we did not monitor soil moisture at the fine spatial and temporal resolution necessary to test this hypothesis.

Evidence for weak interspecific interactions and limited competitive release carries interesting implications. If most interspecific interactions are weak, then single-species models may be adequate for forecasting population responses to environmental change \citep{adler_forecasting_2012,kleinhesselink_indirect_2015}. Similarly, differences among species in relative abundance should reflect variation in the strength of intraspecific density dependence more than variation in the strength of interspecific interactions \citep{yenni_strong_2012}. A recent global, grassland study showing little variation in the abundance of invasive species in native and invaded ranges on different continents  \citep{firn_abundance_2011} is consistent with the idea that intraspecific interactions are the primary determinant of species abundances. In systems where this is true, effective management and conservation requires understanding the sources and species-specificity of intraspecific density dependence. In fact, an experiment to test our model's skill at predicting release from intraspecific competition could be more important than the test of interspecific competitive release described here.

@@

%\newpage
\bibliographystyle{Ecology}
\bibliography{IPMrefs}
